---
title: "EC 380: Lecture 17"
subtitle: "Global Finance: Exchange Rate Policies"
author: "Philip Economides"
date: "Fall 2022"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, viridis, knitr, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, magrittr, kableExtra, ggtext, data.table)

red_pink <- "#e64173"
met_slate <- "#272822" 
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#2b59c3"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
# Notes directory
dir_slides <- "~/Lectures/01-Intro"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  warning = F,
  message = F
)  
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(crayon.enabled = F)
options(knitr.table.format = "html")
```

```{css, echo = F}
@media print {
  .has-continuation {
    display: block !important;
  }
}


.pull-lefter {
  float: left;
  width: 67%;
}
.pull-rightish {
  float: right;
  width: 25%;
}

.pull-rightish ~ p {
  clear: both;
}

```


# Prologue

---

# Recap


* FX market mechanisms in the medium run driven by business cycles

* Short run variation in exchange rate attributed to monetary policy and speculation

* Parity relationships allow us to identify breakeven points at which investment decisions are made

### Today

* ExR Systems and single currency areas


---


# Topics

<br>

* Reasons for holding foreign reserves, main institutions


* Effect of $\Delta S, \Delta D$ of foreign currency on home currency


* Identify short, medium and long term forces that affect currency value


* __Three rules of gold standard__


* Compare and contrast various exchange rate systems


* Price changes and real exchange rate interactions


* List conditions necessary to form single currency area

---

# Fixed Exchange Rates 

Several possibilities for setting the value of the country’s currency. 

--

One option is to keep the exchange rate constant, or __fixed__. 

--

Also called .hi-pink[pegged exchange rate] systems.

--

Small countries often give up control over their currency and adopt the currency of another country, usually the dollar or the euro. 

--

More commonly, the value of a nation’s money is set equal to a fixed amount of another country’s currency, or less commonly to a basket of several currencies. 

--

If the exchange rate is not allowed to vary, then it is called a .hi-pink[hard peg]. 

---

# Fixed Exchange Rates

<br>

Fixed exchange rates that fluctuate within a set band are .hi-pink[soft pegs] and these, in turn, can take several forms depending on the amount of variation allowed.

--

```{r, echo=FALSE}

tibble(
  `Currency Regime` = c("Hard pegs", "Soft pegs", "Managed floating", 
                        "Independently floating", "Other", "Total"),
  Countries = c(24,88,35,30,13,190)
) %>% kbl() %>% kable_styling(font_size=18)

```

---

# Gold Standard

<br>

Historically, fixed exchange rates were the norm, often within framework that defined the value of a country’s currency in terms of a .hi-pink[fixed amount of gold]. 

--

After WWII, many nations shifted away from gold and pegged the value of their currencies to the USD.

--

During 1970s, use of .hi-pink[flexible exchange rate systems] increased, mostly in high-income industrial economies.

--

Many developing countries adopted flexible systems in the 1980s and 1990s, easily managed with advent of the computer and internet. 

---

# Gold Standard

<br>

There is no best exchange rate system. 

--

Currently, less than half of the world’s nations have flexible exchange rates.

--

<br>

.hi-pink[Gold standards] are one type of fixed exchange rate. 

--

Abandoned nearly everywhere in 1930s during the Great Depression.

--

Restored in a modified form after WWII but has completely disappeared since the 1970s. 


---

# Gold Standard

<br>

Economists overwhelmingly opposed to return of gold standard, first countries to end gold standard were first ones to escape .hi-pink[Great Depression]. 

--

After WWII, Western economies adopted a modified gold standard under the .hi-pink[Bretton-Woods exchange rate system] (1947–1971).

--

<br>

Abandoned in the early 1970s. Gold standards highlight a pure form of fixed exchange rate with a hard peg.


--

Under pure gold standard, nations keep .hi-pink[gold as international reserve]. 

---

# Gold Standard

Gold used to settle most international obligations, and nations must be prepared to trade it for their own currency whenever foreigners attempt to “redeem” the home currency.

--

In this sense, the nation’s money is backed by gold.

--

There are essentially .hi-pink[three rules] that countries must follow in order to maintain a gold exchange standard.

--

* Must fix value of their currency unit in terms of gold.

* Keep supply of domestic money fixed in some constant proportion to their supply of gold. 

* Nations must stand ready and willing to provide gold in exchange for their home country currency

---

# Gold Standard

<br>

If a country _decided to print large quantities of money_ for which there is no gold backing.

--


* Purchases of domestically produced goods would rise.
 
* Would cause domestic prices to rise leading to foreign goods becoming more attractive. 

--
 
As imports in the home country increase, foreigners accumulate an unwanted supply of the home country’s currency. 

--

This is the point at which gold standard would begin to become unhinged. 


---

# Gold Standard

<br>

If gold supplies low relative to supply of domestic currency, gold reserves will begin to run out at some point. 

--

This spells crisis and a .hi-pink[possible end to the gold standard].

--

Under a .hi-pink[fixed exchange rate system], the national supply and demand for foreign currencies may vary but the nominal exchange rate cannot.

--

It is the responsibility of the monetary authorities to keep the exchange rate fixed. 

---

# Fixed ExR Policy Action


Suppose US & UK both on gold standard and the US demand for GBP $\uparrow$.

--

```{r, echo=FALSE, out.width="85%", warning=FALSE}

tibble(
    country  = c("S1", "S1",
                 "D1", "D1",
                 "D2", "D2"),
    pies     = c(10, 90, 90, 5, 90, 5),
    ExR  = c(0.95, 1.40, 1.05, 1.4, 1.15, 1.5)
) %>% ggplot() +
    geom_line(aes(x=pies, y=ExR, color=country), size=2) +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0.8,1.6)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,100)) +
    #coord_fixed() + 
    ggtitle("US market for GBP") + 
    theme(title = element_text(size=14),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=16),
          axis.title.x = element_text(size=16),
          axis.title.y = element_text(size=16),
          axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))+
    geom_point(aes(x=54,y=1.2), size=3) + xlab("Units of GBP") + ylab("ExR, USD/GBP") +
    geom_point(aes(x=64.5,y=1.26), size=3) +
    geom_segment(aes(x=0, xend=54, y=1.2, yend=1.2), linetype="dashed")+
    geom_segment(aes(x=54, xend=54, y=0.8, yend=1.2), linetype="dashed")+
    geom_segment(aes(x=0, xend=64.5, y=1.26, yend=1.26), linetype="dashed")+
    geom_segment(aes(x=64.5, xend=64.5, y=0.8, yend=1.26), linetype="dashed")+
    geom_label(aes(x=59, y=1.17), label="E", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=71, y=1.25), label="E'", size=4, position = position_dodge(0.9),
               vjust = 0) +
  scale_color_manual(values=c("#006797","#0087c5","#cd2828"))

```



---

# Fixed ExR Policy Action

Policymakers must counter weakening dollar and keep $R_{\text{USD per GBP}}$ fixed.

```{r, echo=FALSE, out.width="85%", warning=FALSE}

tibble(
    country  = c("S1", "S1",
                 "D1", "D1",
                 "D2", "D2",
                 "S2", "S2"),
    pies     = c(10, 90, 90, 5, 90, 5, 10, 90),
    ExR  = c(0.95, 1.40, 1.05, 1.4, 1.15, 1.5, .82, 1.27)
) %>% ggplot() +
    geom_line(aes(x=pies, y=ExR, color=country), size=2) +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0.8,1.6)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,100)) +
    #coord_fixed() + 
    ggtitle("US market for GBP") + 
    theme(title = element_text(size=14),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=16),
          axis.title.x = element_text(size=16),
          axis.title.y = element_text(size=16),
          axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))+
    geom_point(aes(x=54,y=1.2), size=3) + xlab("Units of GBP") + ylab("ExR, USD/GBP") +
    geom_point(aes(x=64.5,y=1.26), size=3) +
    geom_point(aes(x=78,  y=1.2), size=3) +
    geom_segment(aes(x=0, xend=54, y=1.2, yend=1.2), linetype="dashed")+
    geom_segment(aes(x=54, xend=54, y=0.8, yend=1.2), linetype="dashed")+
    geom_segment(aes(x=0, xend=64.5, y=1.26, yend=1.26), linetype="dashed")+
    geom_segment(aes(x=64.5, xend=64.5, y=0.8, yend=1.26), linetype="dashed")+
    geom_segment(aes(x=0, xend=78, y=1.2, yend=1.2), linetype="dashed")+
    geom_segment(aes(x=79, xend=78, y=0.8, yend=1.2), linetype="dashed")+
    geom_label(aes(x=59, y=1.17), label="E", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=71, y=1.25), label="E'", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=80, y=1.17), label="E''", size=4, position = position_dodge(0.9),
               vjust = 0) +
  scale_color_manual(values=c("#006797","#0087c5","#cd2828","#cd2828"))

```

---

# Fixed ExR Policy Action

<br>

There are two possibilities for the home country as it sells its gold reserves. 

* Demand for gold satisfied and pressure on its currency eases

* Begins to run out of gold (coffers too light). 

<br>

--

If the latter happens, the home country may be forced into a devaluation that is accomplished by changing the gold price of its currency. 

---

# Fixed ExR Policy Action

<br>

.hi-pink[Pegged exchange rate systems] operate similarly to gold standard except that instead of gold, another currency is used to “anchor” home currency.

--

* Exposed to policies of external third party. 

* Thailand pegged its currency to USD for many years. 

* If Japan and China devalue against the dollar, also devalued against Thai currency, and Thai exports suddenly less competitive. 

--

The simplest way to avoid is peg to a group of currencies. Reduces the importance of any single country’s currency. 

---

# Fixed ExR Policy Action

<br>

Countries that use a .hi-pink[crawling peg] may peg to a single currency or a basket of currencies, but they regularly adjust the exchange rate. 

--

The changes in the peg may occur on a daily basis by a predetermined amount. 

--

May prevent some country-specific problems with a fixed rate but difficult to manage and has been associated with numerous exchange rate crises.

---

# Topics

<br>

* Reasons for holding foreign reserves, main institutions


* Effect of $\Delta S, \Delta D$ of foreign currency on home currency


* Identify short, medium and long term forces that affect currency value


* Three rules of gold standard


* Compare and contrast various exchange rate systems


* __Price changes and real exchange rate interactions__


* List conditions necessary to form single currency area

---

# Real Exchange Rate

<br>

ExR used so far does not really tell us what a foreign currency is worth. 

--

We still do not know the .hi-pink[purchasing power] of our domestic money when it is converted to a foreign currency. 

--

Suppose USD–Malaysian ringgit exchange rate is 0.25 USD per ringgit, constant over the year. 

--

Suppose Malaysian inflation is __4%__ and US inflation __1%__. 

--

After one year, 4 ringgits that cost one dollar will buy __3% less in Malaysia than the dollar buys in the United States__. 

---



# Real Exchange Rate


Higher inflation in Malaysia erodes the value more rapidly than at home. 

--

Converted to ringgits, .hi-pink[real purchasing power] of the dollar has declined even though the exchange rate is still 0.25 USD per ringgit.

--

For foreign exchange participants, key item of interest is .hi-pink[purchasing power] upon conversion, not nominal units of a foreign currency. 

--

US importer deciding between Malaysian and Chinese textiles does not really care if she gets four ringgits per dollar or eight Chinese yuan per dollar.

--

_Volume of textiles that can be purchased in Malaysia with four ringgits and in China with eight yuan is the only thing that matters._


---

# Real Exchange Rate

<br>

.hi-pink[Real ExR] is market exchange rate adjusted for price differences. 

--

Let’s consider case of a cheese vendor who is trying to decide whether to stock her shop with Italian or  locally produced cheese. 

--


Suppose say that a wheel Italian cheese of a given quality costs __200 EUR__ and American cheese of the same quality costs __180 USD__. 

--

Suppose that the nominal exchange rate is __1.20 USD per euro__ so that __180 USD__ is equivalent to __150 EUR__. 

--

Italian cheese costs one-third more than American cheese, .hi-pink[real ExR] is __1.33 wheels__ of American cheese per wheel of Italian cheese.


---

# Real Exchange Rate

<br>

Algebra:

$$\text{Real ExR} \\
=[(\text{R} \times \text{Foreign Price})]/ (\text{Domestic Price})\\
=[(1.20 \text{ per EUR}) \times (200\text{ EUR per wheel})]/( 180 \text{ USD per wheel})\\
=\frac{(240 \text{ USD per wheel of Italian Cheese})}{(180 \text{ USD per wheel of American cheese})}\\
= 1 \frac{1}{3} \text{ wheels of American cheese per wheel of Italian cheese}$$


--

.hi-pink[Whether wine or cheese, process is the same, as you'll see in your homework.]

---

# Real Exchange Rate

<br>

$$R_r = R_n (P^*/P)$$

--

The real rate equals the nominal rate when the price levels (purchasing powers) are the same in both countries.

--


For example, consider the current USD-EUR exchange rate, along with their price levels: $R_r = (1 \text{ USD per EUR}) \times (109.1/108.5)$ 


--

Even though we express parity in EUR to USD, .hi-pink[real exchange rates] suggest appreciated EUR value and slightly greater buying power with the dollar due to lower inflation.  


---

# Choosing Exchange Rate System

<br>

Active area of economic research has focused on performance characteristics of systems under different economic conditions and institutional arrangements. 

--

Researchers have become more concerned with understanding how varying degrees of .hi-pink[flexibility] or .hi-pink[fixity] might best serve the interests of individual countries. 

--

How do different systems influence core elements of a country’s macroeconomy?

---

# Choosing Exchange Rate System

<br>

.hi-pink[Traditional view]: Countries with fixed exchange rate systems were better at controlling inflation but paid a price in the form of slower economic growth. 

--

Governments have to be very careful about issuing new money. Exchange rate policy that limits the supply of money would also help avoid inflation. 

--

Limits placed on ability to manipulate money supply also remove an important monetary policy tool tused to manage economic growth.


---

# Choosing Exchange Rate System

<br>

.hi-pink[Research has failed to demonstrate a strong relationship between the type of exchange rate system and either inflation or economic growth.]

--

Before 90s, countries .hi-pink[pegged exchange rates] associated with lower inflation, but during the 90s the differences disappeared. 

--

Countries with .hi-pink[flexible rates] have higher average rates of economic growth, but result depends on classification of the fastest growing Asian economies. 

--

When they omit countries who manage flexible rates, no significant difference in growth between countries with relatively fixed and relatively flexible rates. 


---

# Choosing Exchange Rate System

Neither fixed nor flexible rates offer superior protection against a currency crisis. 

--

_No particular system_ seems to rank above any other in its ability to provide superior macroeconomic performance.

--

If the goal set to minimize negative shocks to an economy, then .hi-pink[source of the shock] determines which system should be adopted. 

--

If originates in monetary sector (printing new money), fixed rate is better since it imposes discipline on the central bank. 

--

If originates in the external environment (oil shock), relatively more flexibility enables the country to adapt more easily.

---

# Topics

<br>

* Reasons for holding foreign reserves, main institutions


* Effect of $\Delta S, \Delta D$ of foreign currency on home currency


* Identify short, medium and long term forces that affect currency value


* Three rules of gold standar


* Compare and contrast various exchange rate systems


* Price changes and real exchange rate interactions


* __List conditions necessary to form single currency area__


---

# Single Currency Areas

<br>

Jan-1999, 11 members of EU .hi-pink[adopted euro] as their official currency. 

--

As the EU added new members, several chose to use the euro, and as of 2016, .hi-pink[19 of 28 members] had replaced their national currencies with the euro. 

--

Shared vision developed economic interdependence and prevent future war. 

--

Nation’s currency is one of its strongest symbols of national sovereignty, remarkable set of events.

---

# Single Currency Areas

<br>


Four potential reasons why countries might want to share a common currency. 

--

* Eliminates need to convert money, reduces transaction costs. Some gain in efficiency. 


* Eliminates price fluctuations caused by changes in the exchange rate.

* Can help increase political trust between countries seeking to increase their integration. 

--

Nations that give up their national money do not do so without cost. 

---

# Single Currency Areas

Adoption of a .hi-pink[common currency] also means that the country no longer has its own .hi-pink[money supply as a tool] for managing its economic growth. 

--

* If control money supply, can influence growth of economy in the short run thru change in money supply.

* Hard for monetary policy of single currency to implement interest rates that suit every member state perfectly. 

* Countries give up their ability to alter their exchange rates. 

* Expected to implement policies aimed at pushing down prices and wages inside their countries and do not have the option to devalue currencies to make goods cheaper. 


---

# Single Currency Areas

<br>

.hi-pink[Conditions for adoption of single currency]

--

* Business cycle must be synchronized 

--

* High degree of labor and capital mobility between the member countries

--

* Regional policies capable of addressing the imbalances that may develop

--

* Nations involved must be seeking a level of integration that goes beyond simple free trade




---

exclude: true

```{R generate pdfs, include = F, eval = F}
#remotes::install_github('rstudio/pagedown')
library(pagedown)
pagedown::chrome_print("17-ExR.html", output = "17-ExR.pdf")
```