---
title: "EC 380: Lecture 2"
subtitle: "Trade Theory: The Ricardian Model"
author: "Philip Economides"
date: "Winter 2024"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, viridis, knitr, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, magrittr)
#devtools::install_github("R-CoderDotCom/econocharts")

library(econocharts)
# Define colors
red_pink <- "#e64173"
met_slate <- "#272822" 
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#2b59c3"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
# Notes directory
dir_slides <- "~/Lectures/01-Intro"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  warning = F,
  message = F
)  
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(crayon.enabled = F)
options(knitr.table.format = "html")
```

```{css, echo = F, eval = F}
@media print {
  .has-continuation {
    display: block !important;
  }
}
```

# Prologue

---

# Overview Trade Theory

<br>

.hi-pink[Why does trade occur?] Two strands within the field of international trade:

--

__I. Neoclassical models of trade__ ('old' theories):

- Ricardian model: Technology differences spur trade

- Heckscher-Ohlin model: Resource differences spur trade

--

__II. 'New' Trade Theory__

- Krugman model: Love of variety spurs trade

- Melitz model: Heterogeneous firms drive trade


---

# Today's Class

<br>

We will begin looking into international trade theory, starting with the .hi-pink[Ricardo model].

* Single factor of production (labor)

* Two-country model (home, foreign)

* Technology differences across countries 

--

__To reiterate:__ Always remember to describe each model by three attributes, such as those listed above, and the model's main takeaway(s)!

---

# Snapshot of US imports

<br>

In 2018, the US imported approximately 28M USD in snowboards 

* Almost half of which were imported from China
* Austria, UAE and Taiwan were other notable sources

--

The US represents the largest economy in the world, yet it resorts to imports of goods at a level that far exceeds its exports. 

--

For example, while 99% of shipping containers that enter Los Angeles port are fully loaded with goods, _70% of containers leaving the port are empty_. 

---

# Reasons for Trade

<br>

With all the manufacturing capability in the US, why purchase snowboards from abroad instead of producing them domestically?

--

* Technology differences in each country? 
* Total resources available differ in each country?
* Production cost differences in each country?
* Proximity to countries with more productive labor forces?

--

For now, consider .hi-pink[opportunity cost]. According to the __Ricardian perspective__, the US devotes domestic labor that would otherwise be used to produce these snowboards to .hi-pink[comparatively] more productive tasks. 

---

# Comparative Advantage

<br>

During the 16th and 17th century, the study of economics was in its infancy. Misguided beliefs in public policy were widely held across nations.

--

> Mercantilism: a stockpile of gold and silver was considered an appropriate barometer for the state of the domestic economy. 

> Imports $\implies$ less gold and exports $\implies$ more gold.  Tariffs set high to prevent trade deficits.

--

Ricardo would go on to demonstrate that under .hi-pink[balanced free trade] (no tariffs), trade benefits __every country__. This is achieved when every country exports the goods they have .hi-pink[comparative advantage] in.  


---

# Comparative Advantage Overview

__Ricardo's logic:__

--

* Two countries, Portugal and England

--

* Two goods, wine and cloth

--

* Portugal has absolute advantage in production of both goods

--

* England is particularly bad at making wine, making it _relatively_ good at cloth production

--

England has .hi-pink[comparative advantage] in cloth. Portugal has .hi-pink[comparative advantage] in wine. 

--

Free trade puts both countries in a better state than autarky would.  

---

# Ricardian Model of Trade

--

<br>

We will go with our own version of the model. 

* Two countries, Home and Foreign
* Two goods, apple pie and potatoes
* One factor of production, labor

--

Suppose in our case the home country ends up exporting apple pies and importing potatoes.

--

.hi-pink[What would this imply about which good each country has comparative advantage in?]

---

# Key Variables

<br>

.hi-pink[Marginal Product of Labor]: MPL represents the additional units of a good produced, given a one unit increase in the number of workers assigned to a particular task.

--

$MPL^H_{\text{pie}}=9$ and $MPL^H_{\text{potato}}=7$ represent the marginal change in number of pies and potatoes produced by the home country as the number of workers rises by one unit. 

--

Essentially boils down to how "good" each country is per unit of worker. 

--

$MPL^F_{\text{pie}}=10$ and $MPL^F_{\text{potato}}=12$ for Foreign with .hi-pink[absolute advantage]<br> 

---

# Key Variables

<br>

.hi-pink[Production Possibilities Frontier]: Coordinates represent various production bundles of pies and potatoes for each country is capable of producing, upon using its full labor force $\bar{L}=10$.

--

We can plot these rather easily by finding the .hi-pink[four axes intercepts]. Each country-good marginal product of labor times the workforce size $MPL^{\text{country}}_{\text{good}} * \bar L$

--

In our case this leads to...

--

$$\implies \bar{\text{Q}}^{H}_{pie} = 9*10=90, \ \ \ \ \bar{\text{Q}}^H_{\text{potato}}=7*10=70\\
\implies \bar{\text{Q}}^{F}_{pie} = 10*10=100, \ \ \ \ \bar{\text{Q}}^F_{\text{potato}}=12*10=120$$

---

# PPF Scenario

```{r, echo=FALSE, out.width="95%"}

tibble(
  country  = c("Home", "Home",
               "Foreign", "Foreign"),
  pies     = c(90, 0,  100, 0),
  potatoes  = c(0, 70, 0, 120)
) %>% ggplot() +
  geom_line(aes(x=pies, y=potatoes, color=country), size=2) +
  theme_bw() + 
  scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
  scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
  coord_fixed() + ggtitle("Production Possibility Frontiers, Home and Foreign") + 
  theme(title = element_text(size=14),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size=16),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14))

```

---
class: inverse, middle

# Autarky (No Trade)


---

# Autarky (No Trade)

```{r, echo=FALSE, out.width="95%"}

tibble(
  country  = c("Home", "Home"),
  pies     = c(90, 0),
  potatoes  = c(0, 70)
) %>% ggplot() +
  geom_line(aes(x=pies, y=potatoes, color=country), size=2) +
  theme_bw() + 
  scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
  scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
  coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
  theme(title = element_text(size=14),
        legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size=16),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16),
        axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14))

```

---

# Home PPF


__Assumption:__ PPFs straight due to constant MPLs 

__Slope of PPF:__ equal to the marginal rate of substitution between potatoes and pies. 

$$\text{Slope of PPF}^{H} =  - \frac{70}{90} = - \frac{MPL^H_{\text{potato}}*\bar L}{MPL^H_{\text{pie}}*\bar L} = - \frac{MPL^H_{\text{potato}}}{MPL^H_{\text{pie}}} = - \frac{7}{9}$$

--

.hi-pink[Opportunity cost] measured by slope of  PPF. For $\Delta Q^H_{\text{pie}}=1$, this would come at the cost of less potatoes being produced $\implies \Delta Q^H_{\text{potato}} = -\frac{7}{9}$.

--

$\frac{7}{9}$ bags of potatoes is the .hi-pink[opportunity cost] of obtaining 1 more apple pie and the slope of the PPF for Home. 


---

# Indifference Curve

<br>

--

* PPF: whole set of possible production outcomes, all use the entire workforce
* Any production bundle below the PPF line is suboptimal, utilizes $L<\bar L$
* Which of these max bundles do we choose?

--

Depends on home's demand for the two goods. Each .hi-pink[indifference curve] shows consumption bundles of goods that economy can consume and be equally satisfied.

--

The consumer is __indifferent__ in its own preferences across the consumption bundles listed on a particular indifference curve. 

---

# Indifference Curve

```{r, echo=FALSE, out.width="85%"}


p = indifference(ncurves = 2,  # Two curves
                  x = c(2, 4),  # Intersections
                  xlab = "pies",              # X-axis label
                  ylab = "potatoes",              # Y-axis label
                  linecol = 2,  # Color of the curves
                  pointcol = 1) # Color of the intersection points


int <- bind_rows(curve_intersect(data.frame(x = 1:1000, y = rep(3, nrow(p$curve))), p$curve + 1))

p$p + geom_point(data = int, size = 3, color = 1) +
      annotate(geom = "text", x = int$x + 0.25, y = int$y + 0.25, label = "C") +
  theme(axis.text.x = element_text(size=14),
        axis.text.y = element_text(size=14),
        axis.title.x = element_text(size=16),
        axis.title.y = element_text(size=16))

```

--

Indifferent between A and B and prefers any point on $I_2$ to $I_1$.


---

#  Home Equilibrium

<br>

Suppose we are in .hi-pink[closed economy] where Home PPF acts as a budget constraint.

IC of Home's consumers tangent to PPF at __(36, 42)__.

Home consumes 36 pies and 42 potatoes under autarky. 

--

Underlying assumption of .hi-pink[perfectly comptetitive markets] such that goods are sold at cost. 

Given these resource constraints, this market generates the highest consumer welfare. 

---

# Home Equilibrium

```{r, echo=FALSE, out.width="95%", warning=FALSE}

tibble(
    country  = c("Home", "Home"),
    pies     = c(90, 0),
    potatoes  = c(0, 70)
) %>% ggplot() +
    geom_line(aes(x=pies, y=potatoes, color=country), size=2) +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
    theme(title = element_text(size=14),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=16),
          axis.title.x = element_text(size=16),
          axis.title.y = element_text(size=16),
          axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))  + 
    geom_function(fun = function(x){180 - 37*log(x+20)}, xlim=c(5,105))+
    geom_function(fun = function(x){180 - 35.5*log(x+12)} , xlim=c(4,140))+
    geom_function(fun = function(x){180 - 30*log(x-20)} , xlim=c(5,145))+
    geom_point(aes(x=36,y=42), size=3) +
    geom_label(aes(x=68,y=42), label="Home No-Trade Equilibrium", size=3, position = position_dodge(0.9),
    vjust = 0)
```


---

# Opportunity Cost and Prices

<br>

* Slope of the PPF reflects the .hi-pink[opportunity cost] of producing one apple pie

--

* Under perfect competition the .hi-pink[opportunity cost] of should also equal the relative price of pie

--

* This follows from the economic principle that price reflects the .hi-pink[opportunity cost] of a good 

--

* We can now check that this equality between the .hi-pink[opportunity cost] and the relative price of wheat holds at point A.

---

# Wages


<br>

How do we solve for prices of potatoes and pies at Home under autarky?

--

* .hi-pink[Perfectly competitive market]: Workers hired until value of additional worker is equal to marginal cost of additional worker
* Value represents the price of the good sold times the marginal change in units of good produced
* Workers are paid the equivalent of their MPL times the price of that additional unit they produce


---

# Wages

<br>

Potato (pie) firms hire until wage equals $P_{\text{potato}} *MPL_{\text{potato}}$ $(P_{\text{pie}} *MPL_{\text{pie}})$

--

If we assume labor moves freely between markets, we can assume wages equalized across these industries. 

--

$$P_{\text{potato}} *MPL_{\text{potato}} = P_{\text{pie}} *MPL_{\text{pie}}$$
--

We can rearrange such that the price ratio is equal to the MPL ratio


$$\frac{P_{\text{pie}}}{P_{\text{potato}}} = \frac{MPL_{\text{potato}}}{MPL_{\text{pie}}}$$

Relative price of pies is equal to the opportunity cost of pies at the .hi-pink[no-trade equilibrium]. 


---
class: inverse, middle

# Foreign Country

---

# Foreign Country

Introduce Foreign Country into the model.

Due to broadly superior technology, labor achieves higher MPL for both goods: 

--

$MPL^H_{\text{pie}}=9$ and $MPL^H_{\text{potato}}=7$ represent the marginal change in number of pies and potatoes produced by the home country as the number of workers rises by one unit. 

--

$MPL^F_{\text{pie}}=10$ and $MPL^F_{\text{potato}}=12$ for Foreign with .hi-pink[absolute advantage]

--

Home will still export to Foreign. Why? .hi-pink[Comparative Advantange].   

---

# Comparative Advantage

<br>

What is the .hi-pink[opportunity cost] of producing one additional apple pie? This should be equal to the relative price of pies.

> Home: $\frac{P^H_{\text{pie}}}{P^H_{\text{potato}}} = \frac{MPL^H_{\text{potato}}}{MPL^H_{\text{pie}}} = \frac{7}{9}$

> Foreign: $\frac{P^F_{\text{pie}}}{P^F_{\text{potato}}} = \frac{MPL^F_{\text{potato}}}{MPL^F_{\text{pie}}} = \frac{12}{10}$

Home has a .hi-pink[comparative advantange] in producing pies because home's opportunity cost of producing an additional pie is lower than Foreign's. 

---

# PPF: Foreign Country

```{r, echo=FALSE, out.width="95%", warning=FALSE}

tibble(
    country  = c(
        "Foreign", "Foreign"),
    pies     = c(100, 0),
    potatoes  = c(0, 120)
) %>% ggplot() +
    geom_line(aes(x=pies, y=potatoes, color=country), size=2) +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Foreign") + 
    theme(title = element_text(size=14),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=16),
          axis.title.x = element_text(size=16),
          axis.title.y = element_text(size=16),
          axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))  + 
   # geom_function(fun = function(x){180 - 37*log(x+20)}, xlim=c(5,105))+
   #    geom_function(fun = function(x){180 - 35.5*log(x+12)} , xlim=c(4,140))+
    geom_function(fun = function(x){180 - 30*log(x-6)} , xlim=c(5,145))+
    geom_point(aes(x=30,y=84), size=3) +
    geom_label(aes(x=60,y=87), label="Foreign No-Trade Equilibrium", size=3, position = position_dodge(0.9),
    vjust = 0)


```


---

# Next Time

Currently, we have kept both countries closed off. 

These countries only consume the bundle of goods that they produce. 

--

From autarky scenario $\implies$ free trade. How will .hi-pink[equilibrium outcomes] change?

* What will they produce?

* How much will they consume?

* Do prices change?

* Who benefits from trade under our assumptions?

__Readings:__ International Economics, Ch.3.1 to Ch.3.3


---

exclude: true

```{R generate pdfs, include = F, eval = F}
#remotes::install_github('rstudio/pagedown')
library(pagedown)
pagedown::chrome_print("02-ricardian.html", output = "02-ricardian.pdf")
```
