---
title: "EC 380: Lecture 4"
subtitle: "Trade Theory: Heckscher-Ohlin (HO) Model"
author: "Philip Economides"
date: "Fall 2022"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, viridis, knitr, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, magrittr, kableExtra)

red_pink <- "#e64173"
met_slate <- "#272822" 
turquoise <- "#20B2AA"
orange <- "#FFA500"
red <- "#fb6107"
blue <- "#2b59c3"
green <- "#8bb174"
grey_light <- "grey70"
grey_mid <- "grey50"
grey_dark <- "grey20"
purple <- "#6A5ACD"
slate <- "#314f4f"
# Notes directory
dir_slides <- "~/Lectures/01-Intro"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  warning = F,
  message = F
)  
opts_chunk$set(dev = "svg")
options(device = function(file, width, height) {
  svg(tempfile(), width = width, height = height)
})
options(crayon.enabled = F)
options(knitr.table.format = "html")
```

```{css, echo = F, eval = F}
@media print {
  .has-continuation {
    display: block !important;
  }
}
```

# Prologue

---

# To Recap

### Last Time

* Countries with no absolute advantage in production can still trade!

--

* Since then, the .hi-pink[internet] and subsequent .hi-pink[information age] have largely narrowed differences in technology

--

* .hi-pink[What else helps us understand what drives our need to trade goods?]

--

### Today


.hi-slate[Hechscher-Ohlin Model] suggests differences in factor endowments can explain trade patterns


---

# Hechscher-Ohlin Model

<br>

--

We will go with our own version of the model. 

* Two countries, Home and Foreign
* Two goods, steel and cloth
* .hi-pink[Two] factors of production, labor and .hi-pink[capital]

--

Suppose in our case the home country ends up exporting apple pies and importing potatoes.

--

.hi-pink[What would this imply about which good each country has comparative advantage in?]


---

# Hechscher-Ohlin Model

--

__Key terms:__

--

.hi-slate[Factor abundance]: Relative measure, where if particular factor $x$ represents a large share of total factors, country is $x$-factor abundant. 

.hi-slate[Factor scarcity]: Country has less of a particular factor relative to other factors or in comparison to another country. 

--

The .hi-pink[capital-labor ratio] is a comment tool of comparison for assigning countries into resource endowment groups. 

--

The higher $K/L$ is for a given country, relative to others, the more capital-abundant it is. 

---

# Hechscher-Ohlin Model

<br>

```{r, echo=FALSE}

tibble(
  x = c("Capital", "Labor"),
  y = c("90 Tractors", "300 Farmers"),
  z = c("150 Tractors", "400 Farmers")
) %>% kbl(col.names=c("", "Ireland", "United Kingdom")) %>%
  kable_styling()

```

--

<br>

.hi-pink[Which country is capital-abundant?]

--

$$\frac{K_{IRL}}{L_{IRL}} = 0.3, \ \ \ \frac{K_{UK}}{L_{UK}} = 0.375$$


---

# Hechscher-Ohlin Model

### Production Costs

<br>

--

The ratio between capital and labor implies .hi-pink[abundancy]. Countries where a given factor is relatively .hi-pink[more abundant] exhibit .hi-pink[lower input prices] per unit of the factor. 

--

The labor-abundant country finds labor to be relatively cheaper, per unit, than the capital-abundant country.

--

HO: Labor-abundant country has a .hi-pink[comparative advantage] due to its edge the cost of production of .hi-pink[labor-intensive goods]. 

---

# Hechscher-Ohlin Model

### Production Costs

<br>

--

Since the US is .hi-pink[capital-abundant], it faces relatively cheaper capital and faces a lower opportunity cost in production that uses relatively more capital.

--

This may explain US trade patterns in which .hi-pink[capital-intensive] exports of jet engines and agricultural products dominate its goods outflows. 



---

# Gains from Trade

--

<br>

Ricardian model assumed a single factor $\implies$ all countries face same trade-off regardless of input levels.

--

HO model now considers combinations of factors, where some specific combination of two is most productive. 

--

.hi-pink[How does this affect our visualization?]

Production possibilities frontier (PPF) will be curved instead of straight. 

---

# New PPF

```{r, echo=FALSE, out.width="95%", warning=FALSE}

tibble(
    country  = c("Home Pre-Trade", "Home Pre-Trade",
                 "Home Trade", "Home Trade"),
    pies     = c(90, 0, 90, 5),
    potatoes  = c(0, 70, 0, 85)
) %>% ggplot() +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
    ylab("Traditional Textiles Output") + xlab("Manufacturing Output") +
    geom_segment(aes(x=0, xend=10, y=99, yend=99), linetype="dashed")+
    geom_segment(aes(x=10, xend=10, y=0, yend=99), linetype="dashed")+
    geom_segment(aes(x=0, xend=80, y=27, yend=27), linetype="dashed")+
    geom_segment(aes(x=80, xend=80, y=0, yend=27), linetype="dashed")+
    geom_segment(aes(x=0, xend=20, y=97, yend=97), linetype="dashed", color="red")+
    geom_segment(aes(x=20, xend=20, y=0, yend=97), linetype="dashed", color="red")+
    geom_segment(aes(x=0, xend=75, y=40, yend=40), linetype="dashed", color="red")+
    geom_segment(aes(x=75, xend=75, y=0, yend=40), linetype="dashed", color="red")+
    theme(title = element_text(size=14),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=16),
          axis.title.x = element_text(size=16),
          axis.title.y = element_text(size=16),
          axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))  + 
    geom_function(fun = function(x){100 - 0.02*x - 0.003*x^2 - 0.0001*x^3} , xlim=c(0,200))

```

---

# New PPF

<br>

As you can see, adjustments at the tailend of these frontiers require a disproportionately large exchange on equipment.  

--

Opportunity costs are .hi-pink[rising] for each type of production. 

--

Each unit increase in the labor-good leads to an increasingly sizeable loss of the other unit. 

--

.hi-pink[Why?]

As you reallocate resources from a capital-intensive good to labor-intensive, you need greater amounts of factors due to factor combinations being misaligned. 


---

# Gains from Trade: Autarky

--

.hi-slate[Closed economy scenario:] Consider a case in which the US does __not__ trade. 

In absense of trade, Home consumes exactly what it produces. 

--

```{r, echo=FALSE, out.width="75%", warning=FALSE}


tibble(
    country  = c("Home Pre-Trade", "Home Pre-Trade",
                 "Home Trade", "Home Trade"),
    pies     = c(90, 0, 90, 5),
    potatoes  = c(0, 70, 0, 85)
) %>% ggplot() +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
    ylab("Traditional Textiles Output") + xlab("Manufacturing Output") +
    geom_segment(aes(x=0, xend=50, y=80, yend=80), linetype="dashed", color="#0080bb")+
    geom_segment(aes(x=50, xend=50, y=0, yend=80), linetype="dashed", color="#0080bb")+
    geom_point(aes(x=50, y=80), size=4, color="#0080bb")+
    theme(title = element_text(size=17),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=20),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20),
          axis.text.x = element_text(size=17),
          axis.text.y = element_text(size=17))  + 
    geom_function(fun = function(x){100 - 0.02*x - 0.003*x^2 - 0.0001*x^3} , xlim=c(0,200))+
    geom_function(fun = function(x){320 - 60.9*log(x+2)} , xlim=c(4,200))+ 
    geom_function(fun = function(x){320 - 58.9*log(x-10)} , xlim=c(4,200))+
    geom_label(aes(x=46,y=73), label="A", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=30), label="U1", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=45), label="U2", size=4, position = position_dodge(0.9),
               vjust = 0) 

```


---

# Gains from Trade: Free Trade

<br>

--

Upon opening up trade and facing no trade costs, the world PPF is now the item of material interest.

--

The previous relative prices of goods changes to a .hi-pink[world relative price]. This identifies the slope of the CPC described in IE Ch. 4.2.2. 

--

Given that this line is not tangental to the production possibility frontier, there are other bundles of goods that reach higher indifference curves $U_i$. 

--

In our case, Home reaches $U_2$ by reallocating production away from textiles and towards manufacturing, given that Home is a capital intensive nation. 

---

# Gains from Trade: Free Trade

```{r, echo=FALSE, out.width="95%", warning=FALSE}


tibble(
    country  = c("Home Pre-Trade", "Home Pre-Trade",
                 "Home Trade", "Home Trade"),
    pies     = c(90, 0, 90, 5),
    potatoes  = c(0, 70, 0, 85)
) %>% ggplot() +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
    ylab("Traditional Textiles Output") + xlab("Manufacturing Output") +
    geom_segment(aes(x=0, xend=50, y=80, yend=80), linetype="dashed", color="#0080bb")+
    geom_segment(aes(x=50, xend=50, y=0, yend=80), linetype="dashed", color="#0080bb")+
    geom_point(aes(x=50, y=80), size=4, color="#0080bb")+
    geom_point(aes(x=80, y=30), size=4, color="#a42a41")+
    theme(title = element_text(size=17),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=20),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20),
          axis.text.x = element_text(size=17),
          axis.text.y = element_text(size=17))  + 
    geom_function(fun = function(x){100 - 0.02*x - 0.003*x^2 - 0.0001*x^3} , xlim=c(0,200))+
    geom_function(fun = function(x){320 - 60.9*log(x+2)} , xlim=c(4,200))+ 
    geom_function(fun = function(x){320 - 58.9*log(x-10)} , xlim=c(4,200))+
    geom_function(fun = function(x){190 - 2.2*x} , xlim=c(20,80), linetype="dashed", color="red")+
    geom_function(fun = function(x){208 - 2.25*x} , xlim=c(30,90), linetype="dashed", color="red")+
    geom_label(aes(x=46,y=73), label="A", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=75,y=25), label="B", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=30), label="U1", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=45), label="U2", size=4, position = position_dodge(0.9),
               vjust = 0) 

```

---

# Gains from Trade: Free Trade

```{r, echo=FALSE, out.width="95%", warning=FALSE}


tibble(
    country  = c("Home Pre-Trade", "Home Pre-Trade",
                 "Home Trade", "Home Trade"),
    pies     = c(90, 0, 90, 5),
    potatoes  = c(0, 70, 0, 85)
) %>% ggplot() +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
    ylab("Traditional Textiles Output") + xlab("Manufacturing Output") +
    geom_segment(aes(x=0, xend=50, y=80, yend=80), linetype="dashed", color="#0080bb")+
    geom_segment(aes(x=50, xend=50, y=0, yend=80), linetype="dashed", color="#0080bb")+
    geom_point(aes(x=50, y=80), size=4, color="#0080bb")+
    geom_point(aes(x=80, y=30), size=4, color="#a42a41")+
    geom_point(aes(x=37.5, y=125), size=4, color="#a42a41")+
    theme(title = element_text(size=17),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=20),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20),
          axis.text.x = element_text(size=17),
          axis.text.y = element_text(size=17))  + 
    geom_function(fun = function(x){100 - 0.02*x - 0.003*x^2 - 0.0001*x^3} , xlim=c(0,200))+
    geom_function(fun = function(x){320 - 60.9*log(x+2)} , xlim=c(4,200))+ 
    geom_function(fun = function(x){320 - 58.9*log(x-10)} , xlim=c(4,200))+
    geom_function(fun = function(x){190 - 2.2*x} , xlim=c(20,80), linetype="dashed", color="red")+
    geom_function(fun = function(x){208 - 2.25*x} , xlim=c(30,90), linetype="dashed", color="red")+
    geom_label(aes(x=46,y=73), label="A", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=75,y=25), label="B", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=42,y=127), label="C", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=30), label="U1", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=45), label="U2", size=4, position = position_dodge(0.9),
               vjust = 0) 

```

---

# Gains from Trade: Free Trade

```{r, echo=FALSE, out.width="95%", warning=FALSE}


tibble(
    country  = c("Home Pre-Trade", "Home Pre-Trade",
                 "Home Trade", "Home Trade"),
    pies     = c(90, 0, 90, 5),
    potatoes  = c(0, 70, 0, 85)
) %>% ggplot() +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
    ylab("Traditional Textiles Output") + xlab("Manufacturing Output") +
#    geom_segment(aes(x=0, xend=50, y=80, yend=80), linetype="dashed", color="#0080bb")+
#    geom_segment(aes(x=50, xend=50, y=0, yend=80), linetype="dashed", color="#0080bb")+
#    geom_point(aes(x=50, y=80), size=4, color="#0080bb")+
    geom_point(aes(x=80, y=30), size=4, color="#a42a41")+
    geom_point(aes(x=37.5, y=125), size=4, color="#a42a41")+
    theme(title = element_text(size=17),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=20),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20),
          axis.text.x = element_text(size=17),
          axis.text.y = element_text(size=17))  + 
    geom_function(fun = function(x){100 - 0.02*x - 0.003*x^2 - 0.0001*x^3} , xlim=c(0,200))+
    geom_function(fun = function(x){320 - 60.9*log(x+2)} , xlim=c(4,200))+ 
    geom_function(fun = function(x){320 - 58.9*log(x-10)} , xlim=c(4,200))+
    geom_function(fun = function(x){190 - 2.2*x} , xlim=c(20,80), linetype="dashed", color="red")+
    geom_function(fun = function(x){208 - 2.25*x} , xlim=c(30,90), linetype="dashed", color="red")+
 #   geom_label(aes(x=46,y=73), label="A", size=4, position = position_dodge(0.9),
#               vjust = 0) +
    geom_label(aes(x=75,y=25), label="B", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=42,y=127), label="C", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=30), label="U1", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=45), label="U2", size=4, position = position_dodge(0.9),
               vjust = 0) 

```

---


# Gains from Trade: Free Trade

```{r, echo=FALSE, out.width="95%", warning=FALSE}


tibble(
    country  = c("Home Pre-Trade", "Home Pre-Trade",
                 "Home Trade", "Home Trade"),
    pies     = c(90, 0, 90, 5),
    potatoes  = c(0, 70, 0, 85)
) %>% ggplot() +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,150)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,150)) +
    coord_fixed() + ggtitle("Production Possibility Frontier, Home") + 
    ylab("Traditional Textiles Output") + xlab("Manufacturing Output") +
#    geom_segment(aes(x=0, xend=50, y=80, yend=80), linetype="dashed", color="#0080bb")+
#    geom_segment(aes(x=50, xend=50, y=0, yend=80), linetype="dashed", color="#0080bb")+
#    geom_point(aes(x=50, y=80), size=4, color="#0080bb")+
    geom_point(aes(x=80, y=30), size=4, color="#a42a41")+
    geom_segment(aes(x=0, xend=80, y=30, yend=30), linetype="dashed", color="#0080bb")+
    geom_segment(aes(x=80, xend=80, y=0, yend=30), linetype="dashed", color="#0080bb")+
    geom_point(aes(x=37.5, y=125), size=4, color="#a42a41")+
    geom_segment(aes(x=0, xend=37.5, y=125, yend=125), linetype="dashed", color="#0080bb")+
    geom_segment(aes(x=37.5, xend=37.5, y=0, yend=125), linetype="dashed", color="#0080bb")+
    theme(title = element_text(size=17),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=20),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20),
          axis.text.x = element_text(size=17),
          axis.text.y = element_text(size=17))  + 
    geom_function(fun = function(x){100 - 0.02*x - 0.003*x^2 - 0.0001*x^3} , xlim=c(0,200))+
    geom_function(fun = function(x){320 - 60.9*log(x+2)} , xlim=c(4,200))+ 
    geom_function(fun = function(x){320 - 58.9*log(x-10)} , xlim=c(4,200))+
    #geom_function(fun = function(x){190 - 2.2*x} , xlim=c(20,80), linetype="dashed", color="red")+
    geom_function(fun = function(x){208 - 2.25*x} , xlim=c(30,90), linetype="dashed", color="red")+
 #   geom_label(aes(x=46,y=73), label="A", size=4, position = position_dodge(0.9),
#               vjust = 0) +
    geom_label(aes(x=75,y=25), label="B", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=42,y=127), label="C", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=30), label="U1", size=4, position = position_dodge(0.9),
               vjust = 0) +
    geom_label(aes(x=100,y=45), label="U2", size=4, position = position_dodge(0.9),
               vjust = 0) 

```


---

# Gains from Trade: Free Trade

<br>

--

Since we are trading at .hi-pink[world price level], we can reach .hi-pink[higher] indifference curves.

--

We .hi-pink[produce at B] and .hi-pink[consume at C] at Home country.

--

Therefore if we produce more .hi-pink[manufacuturing output] than we consumer, we must be .hi-pink[exporting] a subset of these goods (difference between B outcome and C outcome). 

--

In contrast, we produce less .hi-pink[textiles] than we consume, suggesting that Home is .hi-pink[importing] the difference.

---

# Comparing to Ricardian Model

--

.hi-pink[Why B?]

We must produce where the opportunity cost of producing a manufactured good is equal to the relative world price slope (CPC). 

--

Our notion of trade gains are pretty similar to Ricardian views, but specialization is not complete due to .hi-pink[diminishing marginal productivty] associated with each factor. 

--

.hi-pink[Diminishing marginal productivty]: Holding your other factor fixed, the marginal unit increase of a given factor yields increasingly smaller contributions to overall output. 

---

# Comparing to Ricardian Model

```{r, echo=FALSE, out.width="65%", warning=FALSE}


tibble(
    country  = c("Home Pre-Trade", "Home Pre-Trade",
                 "Home Trade", "Home Trade"),
    pies     = c(90, 0, 90, 5),
    potatoes  = c(0, 70, 0, 85)
) %>% ggplot() +
    theme_bw() + 
    scale_y_continuous(expand=c(0,0), limits=c(0,100)) + 
    scale_x_continuous(expand=c(0,0), limits=c(0,400)) + ggtitle("Diminishing Marginal Labor Product") + 
    ylab("Goods Output") + xlab("Labor Input") +
    geom_function(fun = function(x){5 + 0.53*x -0.0008*x^2} , xlim=c(30,380), color="#ffa600")+
  theme(axis.text.y = element_blank(),
        axis.text.x = element_blank(),
        title = element_text(size=17),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text = element_text(size=20),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20))

```

Imagine having one shovel to share between 5 workers. Sure they can exchange the shovel whenever they get tired, but including 10 more is going to make very little difference to total work output.


---

# Trade & Income Distribution

--

In .hi-pink[Ricardian model] we ruled out harmful effects and suggest everyone benefits in society. 

--

Reallocated workers simply left and shrinking industry for an expanding one and were able to exchange their unchanged labor supply for a .hi-pink[larger bundle of goods]. 

--

The .hi-pink[HO model] takes a more moderate view


--

Rather than capital and labor, consider two labor sectors 

* Skilled

* Unskilled

Industries require different combinations of skilled and unskilled labor


---

# Heterogenous Income Effects

<br>

--

We will go into this further later into the course but for now keep in mind that trade openness can have heterogeneous effects depending on which part of the skill bracket a laborer belongs to. 

--

It can be shown that a systematic relationship exists between endowments of factors for a country and who ends up being these winners and losers.

--

I'll show  you a .hi-pink[theoretical argument] for how this can occur. Later we'll examine .hi-pink[empirical evidence], based on applied econometric analysis. 


---

# Stolper-Samuelson Theorem

<br>

Explains theoretical outcomes of assymetric factor outcomes. 

--

* Income depends on input supplied to value of final product

--

* Wages will vary based on their skill level

--

* Labor input earnings (wages) demand on their demand and supply

--

.hi-pink[Derived demand]: Demand for a good or service that is derived from demand for something else (e.g. demand for labor is based on demand for goods and services). 

---

# Stolper-Samuelson Theorem

<br>

Under high output demand, price is high and inputs used to produce benefit by receiving greater returns on their contribution.

--

Any change that impacts prices will have direct impacts on outcomes.

--

Open trade causes the .hi-pink[export good price] to rise and .hi-pink[import good price] to fall. Demand for each factor readjusts, leading to change in returns to each factor. 

--

Resources leave sector of imported good and enter sector of exported good, which cause changes in demand for each input in Ricardian model.

---

# Stolper-Samuelson Theorem

<br>

If manufacturing good is the export, demand for capital factor rises and demand for labor factor falls. 

--

This implies income for actors used intensively for the import sector fall and rises for the intensive factor of the export sector. 

--

.hi-pink[Stopler-Samuelson Theorem suggests that increase in price of good raises income earned by factors intensively used in its production. Fall in price of goods lowers income of factors used intensively.]

---

# Stolper-Samuelson Theorem

<br>

According to this __theory__, a country with a capital-intensive presence (e.g. USA) will shift away from labor demand.

Capital owners will benefit from trade. Laborers will lose out.

--

Read further into the .hi-pink[magnification effect]. This may come up in the future. 

---


# To Recap

* Countries with no absolute advantage in production can still trade in both Ricardian and HO models

* In HO case, advantages are driven by abundancies in different factors which vary across countries

--

* While Ricardian trade openness is a broad gain for everyone, HO highlights a .hi-pink[reallocation] of resources towards specific industries that may disadvantage the factor belonging predominantly to the now partially import-reliant sector. 

--

### Next Class


* Specific Factors model, Model Empirics & Extensions



---

exclude: true

```{R generate pdfs, include = F, eval = F}
#remotes::install_github('rstudio/pagedown')
library(pagedown)
pagedown::chrome_print("04-heck-ohlin.html", output = "04-heck-ohlin.pdf")
```
